---
title: "Project 1"
author: "Philippe Deslauriers"
date: "2026-01-21"
output: pdf_document
---

```{r setup, include=FALSE}
library(rmarkdown)
library(dplyr)
library(ggplot2)
library(DescTools)
```
## Intro

### Data
```{r}
Wind= read.csv("wind-power-production-us.csv", header = TRUE)
Temp= read.csv("average-monthly-surface-temperature.csv", header = TRUE)
WindTS= ts(data = Wind$wind_united_states, end = 265, frequency = 12 ) #Converting the wind data to a time series
TempTS= ts(data = Temp$Average.surface.temperature, start = 6, frequency = 12) #Converting the temperature data to a time series
```
For this project, both time series will be monthly recordings with a start date of January 2001 to January 2023.

-Graphs and info about Wind data
```{r}
plot.ts(WindTS, main= "Average wind power generation for the USA", xlab= "Time in Months", ylab= "Power generated in thousand megawatthours", col= "blue")
acf(WindTS)
pacf(WindTS)
summary(WindTS)
Mode(WindTS)
```
The first data set we look at is the average power generated by wind turbines in the USA. It is measured in thousand megawatthours or 1000xMWh. 
The data was collected from the Energy Information Administration (eia) of the United States. 

-Graphs and Info about Temp data
```{r}
plot.ts(TempTS, main= "Average surface Temperature of the USA", xlab= "time in Months", ylab= "Temperature in Degrees Celsius", col= "orange")
acf(TempTS)
pacf(TempTS)
summary(TempTS)
Mode(TempTS)
```
The Second data set is the average surface temperature of the entire USA, measured in degrees Celsius. The data was collected by the site Our World in Data.

### Differencing



### Filtering 



### Stationarity





#### Bibliography
